---
title: "Unsupervised Learning"
author: "Duong Trinh"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
  pdf_document: 
    number_sections: true
    extra_dependencies: ["mathtools","bbm"]
fontsize: 10pt
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
```

<!--- For HTML Only --->
`r if (knitr:::is_html_output()) '
$\\newcommand{\\mathbbm}[1]{\\mathbb{#1}}$
'`

# K-means Clustering

We want to partition the data into $K$ clusters such that the total within-cluster variation (WCV), summed over all $K$ clusters, is a small as possible. Formally:

$$
min \left\{ \sum_{k=1}^K WCV (C_k)\right\}
$$

using Euclidean distance:

$$
WCV (C_k) = \frac{1}{n_k} \sum_{i,i'\in C_k}\sum_{j=1}^p (x_{ij} - x_{i'j})^2
$$

Minimizing the objective function is not trivial. There is a simple algorithm available.

  1. Randomly assign each observation to a cluster 1 to K.
  
  2. Compute the cluster centroid.
  
  3. Assign each observations to the cluster whose centroid is closest.
  
  4. Repeat 2.- 3. until cluster assignment doesn’t change. But the algorithm is not guaranteed to find the global minimum. Thus one should check different starting values.

# Lab SGPE

```{r, warning=FALSE, message=FALSE}
load("/Users/duongtrinh/Dropbox/FIELDS/Machine Learning/Courses/[2022] SGPE Summer School/Tutorials/9_Unsupervised/Lab_unsupervised_data.RData")

library(tidyverse)

data.wide <- data %>% 
  pivot_wider(id_cols = c(Country.Name.x, Country.Code),
               values_from = `Indicator Value`, names_from = Indicator)

# remove missing variables
data <- na.omit(data.wide)
```




